---
title: "Week 07 - Map Classification"
author: "Sofie"
date: "19/03/2024"
---

```{r setup, include=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(echo=TRUE,
	             cache=TRUE,
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)
```

##  Learn about Classification methods

This exercise demonstrates some possibilities of classifications in R, based on Ch.9 of [Geocomputation](https://geocompr.robinlovelace.net/adv-map.html)

### Required R libraries

* We will use the sf, raster, and tmap packages.
* Additionally, we will use the spData and spDataLarge packages that provide new datasets.

```{r}
pacman::p_load(sf, raster, tmap)

```


### Data sets 

We will the two familiar data sets:  `munic` and `viking` for the entirety of Denmark. 

* Recreate them from .rds files in your data/ folder and convert to sf objects in 25832 crs.
* Calculate the density of viking monuments per municipality by either
-- getting municipality names into the viking dataset, grouping and tallying the viking points by municipality and then joining to municipalities OR
-- doing a spatial join between the two datasets and then grouping and tallying by name  

```{r munic-viking}

# Load datasets distribution
munic <- readRDS("../data/gadm36_DNK_2_sp.rds")
viking <- readRDS("../data/viking.rds")

# Convert munic to sf
munic <- st_as_sf(munic)

# Convert sf objects to match (which is at 25832 crs)
munic <- st_transform(munic, crs = st_crs(viking))

# Perform a spatial join (to join the viking monuments with the municipality information)
#viking_munic <- st_join(viking, munic, join = st_within)
viking_munic <-  munic%>%
  st_join(viking)

# Summarize to get the count of Viking monuments per municipality
#viking_munic <- viking_munic %>%
#  group_by(NAME_2) %>%            # NAME_2 = municipality
#  summarise(viking_count = n())   # column with the number of monuments per municipality

viking_munic <- viking_munic%>%
  group_by(NAME_2)%>% #group by municipality 
  tally() #%>%

# Does this answer "Calculate the density of viking monuments per municipality" by either"??

```


### Existing code

Here are some examples of plotting number of viking monuments in Denmark.
Your role is to create a map based on the suggestions below, selecting the most meaningful classification style.

```{r maps}

# pretty style rounds legend numbers. Default style. 
a <- tm_shape(viking_munic) + tm_polygons(col = "n", style = "pretty", n = 5)

# "Jenks" = further smooths over the gaps.Identifies groups of similar values in the data and maximizes the differences between categories
b <- tm_shape(viking_munic) + tm_polygons(col = "n", style = "jenks", n = 6)

# quantile = divides into 5 even groups (ensures same number of observations in each category). But, the bin ranges can vary widely.
c <- tm_shape(viking_munic) + tm_polygons(col = "n", style = "quantile", n = 5)

# Equal interval style divides the distribution into even groups (equal range). Is appropriate for variables with a uniform distrubution
d <- tm_shape(viking_munic) + tm_polygons(col = "n", style = "equal", n = 5)

# Write maps above to objects and plot them side by side with tmap_arrange() for better comparison
tmap_arrange(a, b, c, d)
```


#### Questions

* 1. What are the advantages and disadvantages of each classification method?
* 2. Choose the best classification and create a map with easily legible legend and all other essentials.(Select a suitable color palette from http://colorbrewer2.org/, north arrow, scale, map title, legend title, reasonable number of breaks in the classification )

- They seem very similar, but I could choose the Quantile classification method, as it divides the data into equal-sized groups and therefore might be very good for comparing distributions.
- Include the north arrow, scale, map title, and legend title.
- 5 classes?

* 3. Which method and how many classes did you end up using to display your data? Why did you select that method?
* 4. What principles did you use in arranging the parts of your map layout the way you did and how did you apply these principles?

# Your solution

```{r}
# /Start Code/ #
pdf("municmap.pdf")
tmap_arrange(a,b,c,d)
dev.off()

# /End Code/ #
```




